% ------------------------------------------------------------------------------
% §3 Data — drafted Phase C from:
%   Direction_clean/README.md (reuse map); 00_inventory; 01_outcome findings;
%   2026-07-06 F2 AEMO reconciliation session log; memory: project_aemo_qed_direction_cost
% ------------------------------------------------------------------------------
\section{Data}\label{sec:data}

\subsection{Sources}

All data are AEMO public archives, extracted for January 2022 through December
2024 (36 months) unless noted.

\emph{Bids.} The complete bid ladder for the focal units --- band prices
(\texttt{BIDDAYOFFER}) and band quantities with declared availability
(\texttt{BIDOFFERPERIOD}) --- for every bid version submitted, 1,578,240
unit-interval rows at the five-minute grain. Each interval is resolved to the
bid version in force at dispatch; the resolution joins cleanly, with zero
intervals missing a price ladder.

\emph{Directions.} AEMO's published direction reports, parsed to an event-level
record of 1,638 directions: unit, instruction type (synchronise or remain),
issue and effective times, and cancellation.\footnote{Direction report
timestamps are published in a spreadsheet format whose parsing initially
introduced a systematic $+10$-hour offset. The bug was found, fixed at the
source, and every window-dependent result in this paper is computed on the
corrected times; pre-correction outputs were archived and are cited nowhere.}
Realised directed status is carried to the five-minute grain (375,264 directed
unit-intervals). Direction costs and unit-level compensation dollars are
available from October 2023, when the reporting format changed; before that,
costs exist only as event-level recovery amounts.

\emph{Compensation price.} The monthly $\dt$ series --- the trailing-365-day
90th percentile of South Australian spot prices --- reconstructed from spot
data and verified against AEMO's published values.

\emph{Costs.} Generator short-run marginal cost (SRMC) is built from
engineering primitives: AEMO ISP heat rates, Aurecon variable operating costs, and daily
STTM Adelaide hub gas prices --- roughly \$120/MWh for the Torrens B units at
mid-sample gas prices. Bids never enter the cost measure.\footnote{A
revealed-cost alternative (backing cost out of the units' own offers) was
tested and rejected on both margins: it implies a heat rate of 0.22~GJ/MWh
against an engineering range of 7--11~GJ/MWh, and it barely moves when gas
triples.
The units' offers do not track their costs, which is itself an early sign of
the conduct this paper studies.}

\emph{Market state.} Regional demand, non-synchronous generation, spot
prices, and interconnector flows (Heywood and Murraylink) for the
residual-demand construction of Section~\ref{sec:design}.\footnote{AEMO
tables \texttt{DISPATCHREGIONSUM}, \texttt{DISPATCHPRICE}, and
\texttt{DISPATCHINTERCONNECTORRES}.} Registered capacities are taken from AEMO's
registration tables: 200~MW per Torrens B unit, 478~MW for Pelican Point,
180~MW for Osborne.

\subsection{External validation against AEMO's published series}

The parsed direction record is validated against AEMO's own published
statistics. The share of time South Australia spent under an active direction,
computed from the parsed events, reproduces the corresponding series in AEMO's
\emph{Quarterly Energy Dynamics} to within 0.1 percentage points in six of nine
overlapping quarters (worst deviation 2.3 points) \citep{AEMO2024_qed_q2}.
Quarterly direction costs, summed from the parsed record under the Rules'
recovery-amount definition, match AEMO's published value for the second
quarter of 2023 to machine precision \citep{AEMO2024_qed_q2}. The event record underlying every direction-window calculation in
this paper is, to measurement accuracy, the operator's own.

\subsection{Panels}

Three panels carry the analysis. The \emph{outcome panel} (1,578,240 rows)
holds the withholding measures of Section~\ref{sec:design} for all five units.
The \emph{regression panel} (1,261,576 rows) restricts to the four test units
and adds essentiality, competition, and market-state controls. The \emph{RQ2
matched panel} (140,259 rows in the base sample; 12,513 matched essential
rows, of which 11,120 remain once the June-2022 suspension window is
excluded) is the coarsened-exact-matched
sample described in Section~\ref{sec:design}. All timestamps are handled in
fixed UTC+10 market time; South Australia's civil daylight saving never enters.
Table~\ref{tab:sumstats} reports unit-level summary statistics.

\input{../output/tables/T0_sumstats}
