% ------------------------------------------------------------------------------
% Appendix C — drafted Phase C from findings_task9/10 (+task 7/8 counts via logs);
%   findings_task13_roster.md
% ------------------------------------------------------------------------------
\section{$N-1$ Essentiality and the Registered Day-Level Tests}\label{app:n1}

The main-text essentiality flag $\pex$ is an $N-0$ condition: the requirement
fails \emph{now} without the unit. Day-level conduct tests need more cells
than $N-0$ provides --- only 80 clean essential days exist in three years ---
so the mechanism tests of Section~\ref{sec:mechanism} use a registered
broader flag, $N-1$ essentiality: the system is one credible contingency away
from needing the unit. The construction is a strict superset of $\pex$
(1,290 essential days against 186; 606 clean against 80, a 7.6-fold gain in
cells), built from the same rivals-only inputs, and passes the same four
validity checks, with a monotone gradient of system-stress measures across
the not-essential / $N-1$ / $N-0$ tiers.

The two registered tests reported in Table~\ref{tab:mechanism} were run under
this flag with their interpretations committed in advance. The exit-act
regression (473 clean evening-on-offer days, 96 essential): essential days
\emph{retain} offered evenings ($-0.163$, bootstrap $p=0.017$; three-tier
split $-0.190$ for $N-1$-only days against an exact zero for $N-0$ days). The
floor-pricing test (650 clean days, 115 essential): no essential-day premium
on the offered floor price (+\$45, $p=0.86$), with within-day essential-hour
effects of +\$65 ($p=0.40$) pooled and exactly \$0 for Pelican Point ---
whose offered floor price is a constant $-\$998$ within days. Median offered
floor prices sit at the floor band ($-\$998$ to $-\$1{,}000$) in every cell
of every unit. An earlier $N-0$ suggestion of a +\$2,414 essential-day price
premium (16 clean days, $p=0.17$) does not survive the powered version.

A final registered comparison closes the loop between the roster and the
security standard. On 93.3\% of station-days the minimum-combination standard
asks for zero Torrens units --- rivals' availability covers it. On the 73
days it asked for one or two, the station's declared roster met or exceeded
the ask on all 73. The station never declares itself short of what the
standard requires: whatever else the absence is, it is calibrated to keep the
station continuously directable.
